\section{Implementation}

\subsection{Integration of External Libraries}

Alongside the C standard library headers, the \texttt{gsl\_rng.h} of the GNU Scientific Library (GSL) as well as the two numeric libraries \texttt{cvc\_numerics.h} and \texttt{cvc\_rng.h} were included in the project. 
Of the standard library, the header files \texttt{stdio.h}, \texttt{stdlib.h}, \texttt{tgmath.h}, and \texttt{time.h} were used, the GSL header \texttt{gsl\_rng.h} was only accessed to generate pseudorandom nubers. 

The \texttt{cvc\_numerics.h} of the two own libraries provides mainly numerical methods for differentation, integration or to solve differential equations, while the \texttt{cvc\_rng.h} offers functions requiring PRNGs.
This includes for instance Monte-Carlo integrators or tools helpful to analyse stochastic datasets such as calculating mean or standard deviation of given arrays.


\subsection{Structure and Workflow of the Main Program}

\subsubsection{Generation of Random Numbers by a Static PRNG}

For the generation of pseudorandom numbers, the in \prettyref{ssec:methodology_rng_mt19937} described Mer\-sen\-ne\,Twis\-ter MT19937 of the GSL library was used. It was implemented statically 
to generate uniformly distributed numbers in the interval $[0,\,1)$ and can therefore be accessed globally by the respective functions. 

Overall the PRNG was used in two different ways, once to obtain probabilities, other to generate random integers.
The former was accomplished by checking if the uniform was smaller than the respective probability, the latter by multiplying the uniform by the largest desired integer and then casting it to \texttt{int}.


\subsubsection{Implementation of the Modelling Grid}\label{ssec:impl_modelling_grid}

The above mentioned grid itself was realized as a $\left(L+2\right)\times \left(L+2\right)$ heap section of integer values, with $L$ being the sidelength of the quadratic grid where the actual spread of the
infection takes place. While this section is technically one-dimensional, it will for simplicity reasons be here referred to as a two-dimensional structure of the given shape. 
Inside the grid, the following integer values were used to model the different states of the people within the simulation:
\begin{itemize}
    \item \texttt{0}: this person is Susceptible \susceptible{} to the infection
    \item \texttt{1}: the person is Infected \infected{}
    \item \texttt{2}: the person is Recovered \recovered{} and currently not susceptible
    \item \texttt{-1}: the person is Vaccinated \vaccinated{} and does not participate in the spread
\end{itemize}
The grid was implemented with a supporting edge of ghosts at the top, bottom, left and right border, that are neither infectious nor subject to any updates of the grid --- they will permanently take the value
\texttt{0} and are irrelevant for the later visualization and analysis of the data.


\subsubsection{Functions Acting on the Modelling Grid}\label{ssec:impl_functions_on_grid}

In the main program all changes in the grid are facilitated by functions outside the main one. They entirely work with or change the grid in-place and all take the grid itself as well as its length $L+2$ as arguments.
Inside the functions, the grid is refered to as \texttt{grid} and its length as \texttt{length}, while the above probabilities are passed in form of the double array \texttt{probabilities} having $p_1$, $p_2$, $p_3$, 
and $p_4$ as its entries. Excluding the ones to calculate the infection rate and its time average, none of the functions does have a return value. Overall the following functions can be called from the main:
\begin{itemize}
    \item \texttt{void print\_grid}: prints the passed grid as terminal ouput
    \item \texttt{void grid\_init}: initializes the grid applying the vaccination probability
    \item \texttt{void update\_node}: updates the given node selected by its $L\times L$ grid coordinates \texttt{row\_i} and \texttt{column\_j} according to the turnover probabilities
    \item \texttt{void grid\_update\_linear}: calls the \texttt{update\_node} function for each node in order of their grid position
    \item \texttt{void grid\_update\_stochastic}: calls \texttt{update\_node} for $L^2$ randomly chosen nodes
    \item \texttt{double ratio\_infected}: calculates the ratio of infected individuals with respect to the total grid population for a given grid
    \item \texttt{double average\_ratio\_infected}: applies \texttt{T} simulation steps to the grid by calling the stochastic grid updater with passed turnover probabilities 
    and averages the calculated infection rate over each simulation step
\end{itemize}


\subsubsection{Structure of the Main Function and Generated Data Files}

The main function is divided into compartements for each individual analysis step, where every compartment is a local scope in order to prevent the variables from interfering with each other.
Therefore each one uses its own variables, files and grids, with the latter being then shaped by calling the functions from \prettyref{ssec:impl_functions_on_grid}. The files are also being written at
directly from the main function.

The first compartment itself uses the command line interface to take the probabilities $p_1$, $p_2$, $p_3$, and $p_4$ as well as the grid size $L$ as user input. Further it can be specified how the grid is going to be 
updated. Either all grid nodes are updated sequentially (for user input \texttt{0}) or $L^2$ grid nodes are selected randomly. After initializing the grid, 
individual actualization steps can be performed manually according to the previously selected simulation scheme by pressing \texttt{ENTER} until the loop is quit with \texttt{q} confirmed by \texttt{ENTER}.
While here the mechanism to update the grid is subject to the users input, for the following simulation tasks the model operates using the stochastic updating method for each timestep 
(i.e. $L^2$ random nodes are selected for actualization). It was chosen to avoid adding a spatial bias by having a pre-selected order; however,
in terms of results, there is no meaningful difference expected between the two schemes.

The remainder of the program runs independently from the user, \textbf{overall a running time of around three minutes should be expected after the interactive part is finished}. 
The majority of the time is comsumed by the second scope, generating the data files for the $p_1$ dependency of time-averaged infection rate $\overline{\langle I\rangle}$ without vaccinations.
The files do each have the respective grid size in column one and the values of $p_1$ written in the first row. The data points are the time-averaged infection rates for the previous parameters and calculated
using the corresponding calculation function for $T=1000$. Overall three files are being written at in the \texttt{/soi\_data} directory, one for each combination of $p_2$ and $p_3$:
\begin{itemize}
    \item \texttt{soi\_average\_ratio\_infected\_over\_p1\_a.csv}: $p_2=0.3$, $p_3=0.3$
    \item \texttt{soi\_average\_ratio\_infected\_over\_p1\_b.csv}: $p_2=0.6$, $p_3=0.3$
    \item \texttt{soi\_average\_ratio\_infected\_over\_p1\_c.csv}: $p_2=0.3$, $p_3=0.6$
\end{itemize}

The data file generated by the third compartment for $p_1=p_2=p_3=0.5$,
\begin{itemize}
    \item \texttt{soi\_average\_ratio\_infected\_over\_p4\_v.csv},
\end{itemize}
uses a similar structure. Here the vaccination rate occupies the first row, with the columns once again using the same grid sizes as before.

For the time development of the infection rate, an additional heap section is used to support the statistical analysis across data samples. It has the data series for each time as virtual rows,
that are then passed to the respective \texttt{cvc\_rng.h} functions to calculate mean and standard deviation. The data file of the this third scope,
\begin{itemize}
    \item \texttt{soi\_ratio\_over\_time.csv},
\end{itemize}
has the sample number as fist column and the timestep as first row, the data points are the infection rates of each sample and timestep. 
Mean and standard deviation have extra rows at the end of the file marked with \texttt{00} and \texttt{01} respectively instead of the sample number.

Addinitionally, three data files resembling the state of the grid itself are generated into the \texttt{/soi\_animations} folder:
\begin{itemize}
    \item \texttt{soi\_grid\_over\_time\_a.csv}
    \item \texttt{soi\_grid\_over\_time\_b.csv}
    \item \texttt{soi\_grid\_over\_time\_c.csv}
\end{itemize}

Therefore the dimensionality of the used two-dimensional $96\times 96$ grid is reduced to one and for each timestep occupying the first column of the files, the grid state is a one-dimensional data series behind it.
For the representation of the state of each individual node, the same notation as described in \prettyref{ssec:impl_functions_on_grid} is used.

% Overall, the following files are generated by the main program into the data folder \texttt{/soi\_data}:
% \begin{itemize}
%     \item \texttt{soi\_average\_ratio\_infected\_over\_p1\_a.csv}
%     \item \texttt{soi\_average\_ratio\_infected\_over\_p1\_b.csv}
%     \item \texttt{soi\_average\_ratio\_infected\_over\_p1\_c.csv}
%     \item \texttt{soi\_average\_ratio\_infected\_over\_p4.csv}
%     \item \texttt{soi\_ratio\_over\_time.csv}
% \end{itemize}

% Additionally, the files of the time evolution of the grid states themselves are created in the \texttt{/soi\_animations} folder:
% \begin{itemize}
%     \item \texttt{soi\_grid\_over\_time\_a.csv}
%     \item \texttt{soi\_grid\_over\_time\_b.csv}
%     \item \texttt{soi\_grid\_over\_time\_c.csv}
% \end{itemize}